---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CroPlotR

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Codecov test coverage](https://codecov.io/gh/SticsRPacks/CroPlotR/branch/master/graph/badge.svg)](https://codecov.io/gh/SticsRPacks/CroPlotR?branch=master)
[![R build status](https://github.com/SticsRPacks/CroPlotR/workflows/R-CMD-check/badge.svg)](https://github.com/SticsRPacks/CroPlotR/actions)
<!-- badges: end -->

`CroPlotR` aims at the standardization of the process of analyzing the outputs from crop model such as 
[STICS](https://www6.paca.inrae.fr/stics_eng/), [APSIM](https://www.apsim.info/) or really any model.

> The package is under intensive development and is in a very early version. The functions may heavily change from one version to another until a more stable version is released.


## Table of Contents

* [1. Installation](#1-installation)
* [2. Examples](#2-examples)
  * [2.1 Plotting](#21-plotting)
    * [2.1.1 Dynamic plots](#211-dynamic-plots)
    * [2.1.2 Scatter plots](#212-scatter-plots)
    * [2.1.3 Group comparison](#213-group-comparison)
  * [2.2 Statistics](#22-statistics)
    * [2.2.1 Dynamic plots](#221-simple-case)
    * [2.2.2 Several groups](#222-several-groups)
    * [2.2.3 Statistics plot](#223-statistics-plot)
  * [3. Help](#3-help)
  
## 1. Installation

You can install the released version of CroPlotR from [Github](https://github.com/SticsRPacks/CroPlotR) either using `devtools` or the lightweight `remotes` package:

* With `devtools`

```{r eval=FALSE}
devtools::install_github("SticsRPacks/CroPlotR@*release")
```
* With `remotes`

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("SticsRPacks/CroPlotR@*release")
```

Normaly, all the package dependencies will be installed for CRAN packages.


## 2. Examples

At the moment, only one function is exported for plots [`plot()`](https://sticsrpacks.github.io/CroPlotR/reference/plot.stics_simulation.html) (and its alias `autoplot()`), and one for the statistics [`summary()`](https://sticsrpacks.github.io/CroPlotR/reference/summary.stics_simulation.html). These function should be the only one you need for all your plots and summary statistics.

Here is an example using STICS with a simulation of three situations (called USM in STICS) with their observations: 

- an intercrop of Wheat and pea
- a Pea in sole crop
- a Wheat in sole crop

Let's import the simulation and observation data: 

```{r}
library(CroPlotR)

# Importing an example with three situations with observation:
workspace= system.file(file.path("extdata", "stics_example_1"), package = "CroPlotR")
situations= SticsRFiles::get_usms_list(usm_path = file.path(workspace,"usms.xml"))
sim= SticsRFiles::get_daily_results(workspace = workspace, usm_name = situations)
obs= SticsRFiles::get_obs(workspace =  workspace, usm_name = situations)
```

### 2.1 Plotting

#### 2.1.1 Dynamic plots

Here is an application of dynamic plots for the 3 situations: 

```{r}
plot(sim, obs= obs)
```

Note that the `obs` argument is explicitely named. This is because the first argument of the function is `...` (we'll see why in a minute).

#### 2.1.2 Scatter plots

Here are the same plots, but presented as scatter plots:

```{r}
# Only plotting the first situation for this one:
plots= plot(sim, obs= obs, type = "scatter")
plots[[1]]
```

#### 2.1.3 Group comparison

We can compare groups of simulations alongside by simply adding the simulations objects one after the other (that is why the first argument of the function is `...`). Group simulations can be the results of simulations from different model versions, or simulations with different parameter values.


```{r}
workspace2= system.file(file.path("extdata", "stics_example_2"), package = "CroPlotR")
sim2= SticsRFiles::get_daily_results(workspace = workspace2)

plot(sim, sim2, obs= obs, type = "scatter")
```

Here only one plot is outputed because `workspace2` only contains the intercrop situation.

We can also name the corresponding group in the plot by naming them while passing to the `plot()` function: 

```{r}
plot("New version"= sim, original= sim2, obs= obs, type = "scatter")
```


### 2.2 Statistics

#### 2.2.1 Simple case

Here is an application of summary statistics for the 3 situations: 

```{r}
summary(sim, obs= obs)
```

Note that as for the `plot()` function the `obs` argument is explicitely named. This is because the first argument of the function is `...` to be able to compare groups (i.e. model versions or simulation with different parameter values).

#### 2.2.1 Several groups

We can get statistics for each group of simulations by simply adding the simulations objects one after the other (as for the `plot()` function).

```{r}
summary(sim, sim2, obs= obs)
```

We can also name the corresponding group in the plot by naming them while passing to the `summary()` function: 

```{r}
summary("New version"= sim, original= sim2, obs= obs)
```

By default, all statistics are returned by `summary`, but you can filter them using the `stat` argument:

```{r}
summary("New version"= sim, original= sim2, obs= obs, stat = c("R2","nRMSE"))
```

Please read the help from the [`predictor_assessment()`]((https://sticsrpacks.github.io/CroPlotR/reference/predictor_assessment.html) function.

#### 2.2.3 Statistics plot

It is also possible to plot the statistics: 

```{r}
stats= summary("New version"= sim, original= sim2, obs= obs, stat = c("R2","nRMSE"))
plot(stats)
```

We can choose to plot either the group or the situation in x (and the other is used for grouping and colouring):

```{r}
plot(stats, xvar = "situation", title= "Situation in X")
```


## 3. Help

You can find help for the functions directly using the name of the function followed by the class of the object you need the method for:

- plot:

```{r eval=FALSE}
?plot.stics_simulation

?plot.statistics
```

- statistics:

```{r eval=FALSE}
?summary.stics_simulation
```

As soon as other models are implemented, you'll be able to call their plotting and statistical methods. 


If you have any other problem, please [fill an issue](https://github.com/SticsRPacks/CroPlotR/issues) on Github.


