---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CroPlotR

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Codecov test coverage](https://codecov.io/gh/SticsRPacks/CroPlotR/branch/master/graph/badge.svg)](https://codecov.io/gh/SticsRPacks/CroPlotR?branch=master)
<!-- badges: end -->

`CroPlotR` aims at the standardization of the process of analyzing the outputs from crop model such as 
[STICS](https://www6.paca.inrae.fr/stics_eng/), [APSIM](https://www.apsim.info/) or really any model.

> The package is under intensive development and is in a very early version. The functions may heavily change from one version to another until a more stable version is released.

## Installation

You can install the released version of CroPlotR from [Github](https://github.com/SticsRPacks/CroPlotR) either using `devtools` or the lightweight `remotes` package:

* With `devtools`

```{r eval=FALSE}
devtools::install_github("SticsRPacks/CroPlotR@*release")
```
* With `remotes`

```{r eval=FALSE}
# install.packages("remotes")
remotes::install_github("SticsRPacks/CroPlotR@*release")
```

Normaly, all the package dependencies will be installed for CRAN packages.
