---
title: "List of provided plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{List of provided plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette give a complete list of all the plots that can be created using the package CroPlotR. The plot names given in *italic* correspond to the `type` argument of the respective plot function. This documentation sometimes figures more than one set of essential variables of a plot; in this case any of the sets can be supplied.

```{r, echo = F, message = F}
library(CroPlotR)
library(magrittr)
workspace <- system.file(file.path("extdata", "example_input"), package = "CroPlotR")
soil_data <- readRDS(file.path(workspace, "soil_data_long.rds"))
weather <- readRDS(file.path(workspace, "weather_france.rds"))

soil <- set_soil(
  soil_data, 
  id = "name", 
  layer_thickness = list("epc", "cm"), 
  layer_water_field_cap = list("HCCF", "g/g"),
  layer_water_wilting_pt = list("HMINF", "g/g"), 
  layer_bulk_density = list("DAF", "g/cm^3"),
  organic_N_conc = list("norg", "g/g")
)
```

### Soil

<!-- Sadly, soil_max_wtr_cap doesn't abbreviate to mswc. However, I think it's more important to have the direct relation of `soil_max_wtr_cap` to `layer_max_wtr_cap`. -->
| |   |
|-|---|
|**Type**|*thickness.mswc*|
|**Description**|Thickness against maximal soil water capacity (mswc), the colour shows organic nitrogen concentration.
|**Essential variables**|By soil: `soil_max_wtr_cap`, `thickness` <br/> By layer: `layer_max_wtr_cap`, `layer_thickness` <br/>By layer: `layer_water_field_cap`, `layer_water_wilting_pt`, `layer_water_wilting_pt`, `layer_thickness`|
|**Optional variables**|`organic_N_conc`|
```{r, echo = F, warning = F, fig.align = 'center', fig.width = 7, fig.height = 5} 
plot_soil(soil, type = "thickness.mswc")
```


### Weather
| |   |
|-|---|
|**Type**|*limiting.temperatures*|
|**Description**|Number of hot days against number of cold days. The colour and shape show the year and weather station.
|**Essential variables**|Option 1: `temp_day_max`, `temp_day_min` <br/>Option 2: `temp_day_mean`|
|**Optional variables**|`station_name`, `year`|
```{r, echo = F, fig.align = 'center', fig.width = 7, fig.height = 5} 
plot_weather(weather, type = "limiting.temperatures", size=2)
```

