---
title: "Examples for loading data: STICS and DSSAT"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples for loading data: STICS and DSSAT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introduction
The vignette [Preparing your data for plotting](set_data.html) shows how to transform your data into a usable format for plotting, given that your data is in a suitable format. This vignette gives concrete examples for how to load data from the input files of two major crop models: STICS and DSSAT. The data usied in this vignette is not delivered with the package. To make our code as tracable as possible, we therefore include an illustration of all the data objects used after each chunck of code. (Note that we use a special `head` function for displaying lists that only shows the two first entry of the list and all of it's sub-lists.)
```{r, message = F}
library(CroPlotR)
```
```{r, echo = F}
head.list <- function(obj, n = 2L, ...)
{
    stopifnot(length(n) == 1L)
    origN <- n
    n <- if (n < 0L)
        max(length(obj) + n, 0L)
    else min(n, length(obj))
    lapply(obj[seq_len(n)], function(x)
           {
               tryCatch({
                   head(x, origN, ...)
               }, error = function(e) {
                   x
               })
           })
}
environment(head.list) <- asNamespace('utils')
```

## STICS
### Loading soil data
In order to charge the data, we are going to use the intermediate data object `soil_data` that contains the relevent data before loading it into the `cropr_input` object `soil`.
```{r}
# define path to STICS data files
workspace <- "C:\\Users\\plutz\\Documents\\datasets\\JavaSTICS-1.40-stics-8.50\\example"

# charge soil data from the file `sols.xml` using the SticsRFiles package in long format, 
# add the argument `wide_shape = TRUE` for wide format
soil_data <- SticsRFiles:::get_xml_files_param_df(file_path = file.path(workspace, "sols.xml"))

# create soil object
soil <- set_soil(
  soil_data, 
  id = "name", 
  layer_thickness = list("epc", "cm"), 
  layer_water_field_cap = list("HCCF", "g/g"),
  layer_water_wilting_pt = list("HMINF", "g/g"), 
  layer_bulk_density = list("DAF", "g/cm^3"),
  organic_N_conc = list("norg", "g/g")
)
```
```{r}
head(soil_data)
```

### Loading weather data
Here, we are going to use three intermediate variables before finally creating the `cropr_input` object `weather`:
* usm_list: list of the situations to be simulated, each one corresponding to a single soil x climate x management x cropping system
* meteo_usms: list of the locations of the weather stations
* weather data: data relative to daily measured temperature, rainfall, ...
```{r}
# define path to STICS data files
workspace <- "C:\\Users\\plutz\\Documents\\datasets\\JavaSTICS-1.40-stics-8.50\\example"

# get the list of all simulation unit names
usm_list <- SticsRFiles:::get_usms_list(file.path(workspace,"usms.xml"))

# get the list of all weather file names and name it
meteo_usms <- lapply(usm_list, function(x) 
  unique(unlist(SticsRFiles:::get_param_xml(file.path(workspace,"usms.xml"), 
                                            param_name = c("fclim1", "fclim2"), 
                                            select = "usm", 
                                            value = x)))
  )
names(meteo_usms) <- usm_list

# list of weather data for each simulation unit, 
# every element is a sub-list containing the weather data in one or multiple data.frame(s)
weather_data <- lapply(meteo_usms , function(x) {
  m <- lapply(x, function(y) SticsRFiles:::get_climate_txt(dirpath=workspace, filename = y))
  names(m) <- x 
  m
})

weather <- set_weather(
  weather_data,
  station_name = "station",
  temp_day_max = list("ttmax", "celsius"),
  temp_day_min = list("ttmin", "celsius"),
  year = "year"
)
```
```{r}
head(usm_list)
head(meteo_usms)
# head(weather_data)  # -> produces an error during package build ?
```

## DSSAT

### Loading soil data
```{r, warning = F, message = F}
library(DSSAT)
soil_file <- "C:\\Users\\plutz\\Documents\\DSSAT\\AllSoils\\AU.SOL"

soil_data <- read_sol(soil_file)
soil <- set_soil(soil_data, id="SITE", soil_max_wtr_cap="SSAT")
```

```{r}
head(soil_data)
```

### Loading weather data
```{r, warning = F}
library(DSSAT)
workpath <- "C:\\Users\\plutz\\Documents\\DSSAT\\AllWeather"

weather_names <- list.files(workpath, "\\.WTH$")
names(weather_names) <- weather_names

weather_data <- lapply(weather_names, function(x) read_wth(file.path(workpath, x)))
weather_data <- lapply(weather_data, function(x){ x$year <- format(x$DATE, format = "%Y"); x})

weather <- set_weather(
  weather_data,
  temp_day_max = list("TMAX", "celsius"),
  temp_day_min = list("TMIN", "celsius"),
  year = "year"
)
```
```{r}
head(weather_names)
head(weather_data)
```
